### 2019-06-25 thesisMain.tex

\documentclass[oneside]{scrbook}

\title{A Sample Thesis} 
\author{A.N. Other} 
\date{July 2013} 
\titlehead{A Thesis submitted for the degree of Doctor of Philosophy} 
\publishers{School of Something\\University of Somewhere}

\begin{document} 
\maketitle

\frontmatter 
\tableofcontents 
\listoffigures 
\listoftables

\chapter{Acknowledgements}

I would like to thank my supervisor, Professor Someone. This 
research was funded by the Imaginary Research Council.

\chapter{Abstract}

A brief summary of the project goes here.

\mainmatter

\include{intro}

\backmatter

\end{document}