### 2019-08-06 Titlepage.tex
adjusting header, appendix numbering, title page

\begin{titlepage}
{\fontfamily{phv}\selectfont
\vspace*{-2.9cm}
{\setstretch{0.85}
\centerline{KU Leuven \hfill }
\centerline{Biomedical Sciences Group \hfill }
\centerline{Faculty of Medicine \hfill } 
\centerline{Department of Public Health and Primary Care \hfill } }
\vspace*{-2.4cm}
\hspace{2cm}
\centerline{\phantom{tram 9}\hfill \includegraphics{./logodoctschool.png} }

\begin{minipage}{1.2\textwidth}													%%%
\begin{flushleft} %\large															%%%
%\emph{Autores:}\\	
\vspace{2cm}
\hspace{1cm}\textbf{\Large{DATA SPLITTING AND ITS APPLICATIONS}} \\

\vspace{0.7cm}
\hspace{1cm}\large{Vahid NASSIRI}
\end{flushleft}																	
\end{minipage}

\vspace{2cm}
\begin{minipage}{0.4\textwidth}                                                    %%%
\begin{flushleft}   \small                                                        %%%
\hspace{0.3cm}\underline{Jury}:
\begin{tabular}{llp{0.5cm}l}
                 \vspace{0.5cm}  & \vspace{0.5cm}  &&\\
                 \hspace{0.1cm} Promoter: & Prof. Dr. G. Molenberghs&& Dissertation presented in  \\
                 \hspace{0.1cm} Co-promoter: & Prof. Dr. G. Verbeke&&  partial fulfilment of the \\
                 \hspace{0.1cm} Chair: & Prof. Dr. K. Dierickx&&  requirements for the \\
                 \hspace{0.1cm} Secretary: & Prof. Dr. E. Lesaffre&& degree of Doctor in \\
                 \hspace{0.1cm} Jury members: & Prof. Dr. H. Geys&&  Biomedical Sciences\\
                                                & Prof. Dr. SHelena Geys &&   \\
                                                & Prof. Dr. Dimitris Rizopoulos && \\
                                                &  Prof. Dr. Katrijn Van Deun  && \\
                                                &  Prof. Dr. Marlies Lacante  &&\\
                                                &  Prof. Dr. Karl Meerbergen  &&\\
                                                &  Prof. Dr. Philippe Lemey  &&
               \end{tabular}%
                                                                %%%
\end{flushleft}                                                                     %%%
\end{minipage}                                                   %%%

\vspace{1.2cm}
\begin{center}
\makebox{Date: 18 December 2019}
\end{center}

}

\end{titlepage}
\clearpage